---
title: "Cleaning up iproperty data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning data scraped from iproperty.com.sg

```{r libraries}
library(data.table)
library(ggplot2)
```

Loading data into R using data.table package

```{r loading file}
dat <- fread("iproperty.csv")
```

Cleaning

```{r cleaning}
#remove colon and trailing spaces
names(dat) <- trimws(sub(" :$", "", names(dat)))

# set factor for selected columns
factor.cols <- c("Property Type", "Bedrooms", "Bathrooms", "Tenure", "Estate", "Unit Type", "Carpark", "Air Cond", "District")
for (j in factor.cols) set(dat, j = j, value = as.factor(dat[[j]]))

#combine two name versions taken from different pages
combineTwoColumns <- function (dt, regex) {
  colnames <- dt[, names(.SD), .SDcols = names(dt) %like% regex]
  dt[is.na(get(colnames[1])) | get(colnames[1]) == "View to offer", colnames[1] := get(colnames[2])][, (colnames[2]) := NULL]
}

combineTwoColumns(dat, "^Asking.*(?i)psm")
combineTwoColumns(dat, "^Asking.*(?i)psf")

# set numeric for dollar values
dollar.cols <- dat[, names(.SD), .SDcols = names(dat) %like% "Asking|(?i)psf|(?i)psm" | names(dat) == "Age"]
for (j in dollar.cols) set(dat, j=j, value=(as.numeric(gsub("SGD| |,","",dat[[j]]))))

# convert rows in sq. m. to sq. ft
area.cols = c("Built up", "Land")
for (j in area.cols) dat[, paste0(j,"_sqft") := ifelse(get(j) %like% "sq. m", as.numeric(gsub(" sq. m", "", get(j))) * 10.7639, as.numeric(gsub(" sq. ft.", "", get(j))))]

# Calculate missing Asking Price (PSF)
# note given PSM actually means PSF, but below code uses Asking Price and area to calculate PSF
dat[is.na(`Asking (PSF)`), `Asking (PSF)` := `Asking Price` / `Built up_sqft`]


# Group less frequent property types into "Others"
dist <- as.data.table(table(dat$`Property Type`))
names(dist) <- c("Property Type", "P.Type.Freq")
dist[, property.type.new := ifelse(P.Type.Freq < 200, "Others", `Property Type`)]
dat <- merge(dat, dist, by = "Property Type", all=TRUE)

```

Write cleaned data into CSV

```{r writing to csv}
fwrite(dat, file="iproperty_cleaned.csv")
```

## Plots

Number of listings for each property type
```{r number of property types}
ggplot(dat, aes_string(x="`Property Type`")) + geom_bar() + coord_flip()
```

Types of properties in each district
```{r}
ggplot(dat, aes(x=District, fill=property.type.new)) + geom_bar(position="fill") + coord_flip()
```

Asking Price vs Built Up (sq. ft)
```{r}
ggplot(dat, aes(x=`Built up_sqft`, y=`Asking Price`, color=property.type.new)) + geom_point(alpha=0.1) + xlim(0,10000) + ylim(0,10000000) + stat_smooth(method="lm")

```

Asking Price (PSF) vs Built up by property type
```{r}
ggplot(dat[!is.na(property.type.new) & property.type.new != "Others",], aes(x=`Built up_sqft`, y=`Asking (PSF)`, color=property.type.new)) + geom_point(alpha = .1) + ylim(400,3000) + xlim(0, 2500) + stat_smooth(method="lm")

```

For condominiums, PSF vs built up by district

```{r}
ggplot(dat[property.type.new == "Condominium",], aes(x=`Built up_sqft`, y=`Asking (PSF)`, color=District)) + geom_point(alpha = .5) + ylim(400,3000) + xlim(300, 4000)
```

Density plot of districts
```{r}
ggplot(dat[!is.na(District),], aes(x=`Asking (PSF)`)) + geom_density(aes(fill=District), alpha=0.5) + xlim(0,3000)
```

Density plot of property types
```{r}
ggplot(dat[!is.na(property.type.new)], aes(x=`Asking (PSF)`)) + geom_density(aes(fill=property.type.new), alpha=0.5) + xlim(0,3000)
```
